---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# r4jecs

<!-- badges: start -->
<!-- badges: end -->

`r4jecs` is a small utility package for data analysis workflows used in the
Japan Environment and Children's Study (JECS). The current version focuses on a
small set of reusable helpers for censored environmental data summaries,
significant-figure formatting, and ROPE ranges for regression models.

Older imputation and majority-voting helpers have been retired. For multiple
imputation workflows, use the generic MI/brms pipeline instead:
<https://github.com/fabregithub/generic-mi-brms-pipeline>.

## Installation

You can install the development version of `r4jecs` from GitHub with:

```{r install, eval = FALSE}
# install.packages("devtools")
devtools::install_github("fabregithub/r4jecs")
```

Then load the package:

```{r setup, eval = FALSE}
library(r4jecs)
```

## Main functions

| Function | Purpose |
|---|---|
| `nada_summary()` | Summarise paired measurement/censoring columns using `NADA` |
| `format_sigfig()` | Format numeric values using significant figures |
| `rope_lm()` | Calculate a ROPE range for linear-model coefficients |
| `rope_logit()` | Calculate a ROPE range for logistic-model coefficients |

Backward-compatible wrappers are kept for older scripts:

| Old function | Preferred function |
|---|---|
| `nada.sum()` | `nada_summary()` |
| `nsf()` | `format_sigfig()` |
| `nsf2()` | `format_sigfig()` |

The retired functions `mjvote()` and `tidyMice()` are deprecated and now stop
with a message pointing users to the newer workflow.

## Summarising censored environmental measurements

`nada_summary()` expects alternating columns of measurement values and censoring
indicators:

```{r nada-data, eval = FALSE}
exposure_data <- data.frame(
  lead = c(0.10, 0.20, 0.05, 0.40, 0.12),
  lead_censored = c(FALSE, FALSE, TRUE, FALSE, TRUE),
  cadmium = c(0.01, 0.03, 0.02, 0.04, 0.01),
  cadmium_censored = c(TRUE, FALSE, FALSE, FALSE, TRUE)
)

nada_summary(exposure_data)
```

The expected column order is:

```text
measurement_1, censored_1, measurement_2, censored_2, ...
```

The censoring column should indicate whether each observation is censored, for
example whether the value is below the limit of detection. Rows with missing
measurement or censoring values are removed pairwise so that measurements and
censoring flags remain aligned.

The old function name still works, but new code should use `nada_summary()`:

```{r nada-old-name, eval = FALSE}
nada.sum(exposure_data)
```

## Formatting numbers

Use `format_sigfig()` to format numeric values for tables and reports:

```{r format-sigfig, eval = FALSE}
format_sigfig(c(0.012345, 1.2345, 123.45), sigfig = 3)

format_sigfig(c(0.012345, 1.2345, 123.45), sigfig = 3, digits = 2)
```

The old helpers are retained as wrappers:

```{r format-old-names, eval = FALSE}
nsf(c(0.012345, 1.2345, 123.45), NSF = 3, NBD = 2)
nsf2(c(0.012345, 1.2345, 123.45), digits = 3)
```

## ROPE ranges

`rope_lm()` returns a symmetric ROPE range for linear-model coefficients. By
default, the range is plus or minus 0.1 standard deviations of the outcome.

```{r rope-lm, eval = FALSE}
y <- c(1.2, 2.0, 2.4, 3.1, 4.0)

rope_lm(y)
```

`rope_logit()` returns a symmetric ROPE range on the logit scale. By default,
`change` is interpreted as a relative change around the supplied prevalence.

```{r rope-logit, eval = FALSE}
rope_logit(prevalence = 0.20, change = 0.05)
```

For an absolute probability change instead, use `relative = FALSE`:

```{r rope-logit-absolute, eval = FALSE}
rope_logit(prevalence = 0.20, change = 0.01, relative = FALSE)
```

## Deprecated functions

The following functions are intentionally deprecated:

```{r deprecated, eval = FALSE}
mjvote()
tidyMice()
```

They are kept only to provide a clear error message for older scripts. New work
should use the generic MI/brms pipeline:
<https://github.com/fabregithub/generic-mi-brms-pipeline>.

## Development

After editing roxygen comments or changing package dependencies, regenerate the
package documentation and `NAMESPACE` locally:

```{r dev, eval = FALSE}
devtools::document()
devtools::check()
```

To regenerate `README.md` from this file:

```{r readme, eval = FALSE}
devtools::build_readme()
```
